---
title: "Homework 1 - Group XXX"
output: html_document
date: "2023-11-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Introduction, describe the setting etc.

# Exercise 1: Warm-up & Throwback

Exercise 1 description etc.
BRIEFLY: We have 150 students, *k* of them are liars. Every student must toss a coin *N* times and tell the results obtained. We have to choose *N* and establish a rule to decide whether a student is liar or not (given the *N* tosses).  
Also, we need to design and graphically study a score function which tells how much our strategy is good for every level of *$\alpha$* (probability of claiming that a student is liar given that he's honest, false positive) and *$\beta$* (probability of claiming that a student is honest given that he's a liar, false negative).  
Keep in mind that they depend on *q*, the probability of telling "head" for liars.  
Lastly, add a time constraint: knowing that each flip costs 3 seconds, we want to check at least 75 students in no more than T minutes. We need to tweak the original score function to get a new one that handles this new scenario. Explore this strategy for T=15 minutes and comment findings.

```{r, echo=TRUE}
# R code for exercise 1 (first cell) HERE, echo = TRUE displays code, echo = FALSE is used when displaying plots
# Setting number of students
n <- 150
```

# Exercise 2: Faraway, So Close!

Exercise 2 description etc.  
BRIEFLY: Introduction on p-Wasserstein distance between random vectors and random variables (simpler). In our setting, one random variable represent the **true population** model as a Beta distribution (with arbitrary $\alpha$ and $\beta$ parameters), while the other is an approximation of it: we'll consider a Kernel Density estimator (depending on the random sample extracted from the true population) with bandwidth as the only tunable parameter.   
The objective is to pick an implementation of the kernel density estimator and code its corresponding quantile function. Then, pick two different $(\alpha,\beta)$ pairs and let $\epsilon>0$ be an approximation level to achieve in terms of 1-Wasserstein distance (which is a random variable itself): in other words we're searching for the largest bandwidth such that the 1-Wasserstein distance is $\leq\epsilon$.

```{r, echo=TRUE}
# R code for exercise 2 (first cell) HERE, echo = TRUE displays code, echo = FALSE is used when displaying plots
# PART 0 - SETTING AND PREPROCESSING

# We draw a sample from the true population to obtain a realization of the Kernel Density Estimator
# alpha, beta and n are chosen arbitrarily
alpha <- 0.5
beta <- 2
n <- 1000

# set seed to make the example reproducible
set.seed(5354)
xsample <- rbeta(n, shape1=alpha, shape2=beta)

# set the bandwidth (arbitrary) and obtain a realization of Kernel Density Estimator
h <- 0.1 # this is the actual (arbitrary) bandwidth
ysample_ep <- density(xsample, bw = h/sqrt(5), kernel = "epanechnikov")
ysample_rec <- density(xsample, bw = h/sqrt(12), kernel = "rectangular")

# Riemann sum to integrate the kernel density
F_hr <- function(x, dens, supp, bw){
  return((sum(dens[supp <= x]) * bw)/(sum(dens) * bw))
}

# As the kernel density estimator support falls off the interval [0,1], it is needed to crop and normalize it
# EPANECHNIKOV
C_ep = F_hr(x=1, dens = ysample_ep$y, supp = ysample_ep$x, bw = h) - F_hr(x=0, dens = ysample_ep$y, supp = ysample_ep$x, bw = h) # normalizing constant
ysample_ep$y = ysample_ep$y[ysample_ep$x >= 0 & ysample_ep$x <= 1] # take correct density values
ysample_ep$x = ysample_ep$x[ysample_ep$x >= 0 & ysample_ep$x <= 1] # crop support
ysample_ep$y = ysample_ep$y/C_ep
# RECTANGULAR
C_rec = F_hr(1, dens = ysample_rec$y, supp = ysample_rec$x, bw = h) - F_hr(0, dens = ysample_rec$y, supp = ysample_rec$x, bw = h) # normalizing constant
ysample_rec$y = ysample_rec$y[ysample_rec$x >= 0 & ysample_rec$x <= 1] # take correct density values
ysample_rec$x = ysample_rec$x[ysample_rec$x >= 0 & ysample_rec$x <= 1] # crop support
ysample_rec$y = ysample_rec$y/C_rec
# Epanechnikov kernel is chosen as main ADD TO MARKDOWN PART
```

Here is an example of a random sample extracted from the true population and its respective kernel density estimation:

```{r, echo=FALSE}
par(mfrow=c(1,2))
# Epanechnikov kernel plot
hist(xsample, 
     prob = T,
     main = "Histogram of the random sample",
     sub = "Epanechnikov kernel",
     col = "royalblue",
     xlab = "Value")
lines(ysample_ep,
     lwd = 2,
     col = "darkred")
# Rectangular kernel plot
hist(xsample, 
     prob = T,
     main = "Histogram of the random sample",
     sub = "Rectangular kernel",
     col = "royalblue",
     xlab = "Value")
lines(ysample_rec,
      lwd = 2,
      col = "darkred")
```

WIP - QUANTILE FUNCTION IMPLEMENTATION  
Given a sample (of size $n$) composed by **iid** realizations of a random variable $X$ with PDF (probability density function) $f_X(\cdot)$, the associated kernel density estimator $\hat{f_h}(\cdot)$, parametrized by the bandwidth $h$ and the kernel function $K(\cdot)$, serves as an estimator of $f_X(\cdot)$. It's defined as follows: 

\begin{align}
   \hat{f_h}(x) = \frac{1}{nh} \sum_{i=1}^{n}K\left(\frac{x_i-x}{h}\right)
\end{align}

The kernel function $K(\cdot)$ must be a PDF such that:

- $\int K(x)dx = 1$
- $\int xK(x)dx = 0$ 
- $\int x^2K(x)dx > 0$


If the objective is estimating the CDF (cumulative density function) $F_X(\cdot)$ of $X$, the corresponding estimator $\hat{F_h}(\cdot)$ can be computed from the kernel density estimator $\hat{f_h}(\cdot)$:

\begin{align}
   \hat{F_h}(x) = \int_{-\infty}^{x}\hat{f_h}(t)dt = \frac{1}{n} \sum_{i=1}^{n}G\left(\frac{x-x_i}{h}\right)
\end{align}

Where $G(x) = \int_{-\infty}^{x}K(t)dt$ is the CDF associated to the kernel function.  
Equating this last estimator with $\alpha$ and solving for $x$ gives $\hat{F}_h^{-1}\!(\cdot)$, which is the corresponding estimator of the $\alpha$-quantile:

\begin{align}
   \hat{F}_h^{-1}\!(\alpha) = 
\end{align}

```{r, echo=TRUE}
# PART 1 - COMPUTE CDF^-1 CORRESPONDING TO THE ESTIMATED KERNEL DENSITY
# custom function which computes an approximation of the Kernel Density corresponding CDF
F_h <- Vectorize(function(x, supp, dens){
  f <- approxfun(supp, dens, yleft=0, yright=0)
  cdf <- integrate(f, min(supp), x, subdivisions=1000)$value
  return(cdf)
}, vectorize.args='x')
# Use the uniroot() function to compute the associated quantile function (at level "al")
Q_h <- Vectorize(function(a, s, d){
  tryCatch({
           out = uniroot(function(t) F_h(t, supp = s, dens = d) - a, interval = range(s))$root
           return(out)
  },
         error = function(cond){
           # EDIT MESSAGE
           message("Error with uniroot function in quantile estimation, value replaced with NA")
           return(NA)
  })
}, vectorize.args = 'a')
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
# Estimated vs real CDF (Epanechnikov)
curve(F_h(x, dens = ysample_ep$y, supp= ysample_ep$x), 
      from=0, 
      to=1,
      main = "CDF",
      xlab = "q",
      ylab = expression("P(X"<="q)"),
      sub = "Epanechnikov Kernel")
curve(pbeta(x, shape1=alpha, shape2=beta), 
      from=0, 
      to=1, 
      col="red", 
      add = T)
# Estimated vs real CDF^-1 (Epanechnikov)
curve(Q_h(x, d = ysample_ep$y, s = ysample_ep$x), 
      from=0, 
      to=0.99,
      main = expression(bold("CDF"^-1)),
      xlab = expression("P(X"<="q)"),
      ylab = "q",
      sub = "Epanechnikov Kernel")
curve(qbeta(x, shape1=alpha, shape2=beta), 
      from=0, 
      to=1, 
      col="red", 
      add = T)
```

```{r, echo=FALSE}
# # NOT WORKING WITH RECTANGULAR KERNEL
#  par(mfrow=c(1,2))
#  # Estimated vs real CDF (rectangular)
#  curve(F_h(x, dens = ysample_rec$y, supp= ysample_rec$x), 
#        from=0.15, 
#        to=0.85,
#        main = "CDF",
#        xlab = "q",
#        ylab = expression("P(X"<="q)"),
#        sub = "Rectangular Kernel")
#  curve(pbeta(x, shape1=alpha, shape2=beta), 
#        from=0, 
#        to=1, 
#        col="red",
#        add = T)
#  # Estimated vs real CDF^-1 (rectangular)
#  curve(Q_h(x, d = ysample_rec$y, s = ysample_rec$x), 
#        from=0, 
#        to=0.99,
#        main = expression(bold("CDF"^-1)),
#        xlab = expression("P(X"<="q)"),
#        ylab = "q",
#        sub = "Rectangular Kernel")
#  curve(qbeta(x, shape1=alpha, shape2=beta), 
#        from=0, 
#        to=1, 
#        col="red", 
#        add = T)
```

```{r, echo=FALSE}
# PART 2 - IMPLEMENTATION OF WASSERSTEIN DISTANCE + OPTIMIZATION (Epanechnikov)
# Wrap all up inside one function, to study it at the variation of the bandwidth h
WD <- function(sample, h, ker, al, bet, p){
  # Kernel Density Estimation
  kde <- density(sample, bw = h/sqrt(5), kernel = ker)
  # As the kernel density estimator support falls off the interval [0,1], it is needed    to crop and normalize it
  C = F_hr(x=1, dens = kde$y, supp = kde$x, bw = h) - F_hr(x=0, dens = kde$y, supp = kde$x, bw = h) # normalizing constant
  kde$y = kde$y[kde$x >= 0 & kde$x <= 1] # take correct density values
  kde$x = kde$x[kde$x >= 0 & kde$x <= 1] # crop support
  kde$y = kde$y/C # normalize
  
  # wasserstein distance calculation
  diff <- abs(qbeta(kde$x, shape1=al, shape2=bet)-Q_h(kde$x, s = kde$x, d = kde$y))^p
  f <- approxfun(kde$x, diff, yleft=0, yright=0)
  return((integrate(f, 0, 0.999, subdivisions=1000)$value)^(1/p))
} # Seems to work, comment it

h = 0.1
WD(sample = xsample, h = h, ker = "epanechnikov", al = alpha, bet = beta, p = 1)
```