---
title: "Homework 3 - Group 46"
author: "Name and Surname: Alessandra Colaiocco - ID: 1843714 - Email: colaiocco.1843714@studenti.uniroma1.it \n\n Name and Surname: Gabriele Parisini - ID: 1964381 - Email: parisini.1964381@studenti.uniroma1.it"
output: html_document
date: "2024-01-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Game Theory and Shapley Value
## Introduction
A *game* (in *game theory*) is an abstract mathematical model of a scenario in which *agents* (or *players*) interact. When we're dealing with a game, we're interested in those features of the scenario that are relevant to the decisions of the players.  
In this homework we're focusing on **cooperative games**, which are games populated by a set $\mathcal{P} = \{1,...,p\}$ of players (or *grand coalition*) which are able to form *coalitions* (subsets of the "grand" one).  
To continue further in the explanation, we would need to define the *characteristic function game*:  
  
**Definition 1.**&nbsp;&nbsp;A characteristic function game $G$ is given by a pair $(\mathcal{P},\nu)$, where $\mathcal{P} = \{1,...,p\}$ is a non-empty set and $\nu: 2^{\mathcal{P}}\mapsto\mathbb{R}$ is a **characteristic function** which maps every coalition $C \subseteq\mathcal{P}$ to a real number $\nu(C)$ (which is referred to as the **value of the coalition $C$**).  
  
It's crucial to understand that the characteristic function game assigns a value to a coalition as a whole, **not telling how the coalition value should be divided among the members of the coalition itself.** How to divide the coalition value is a **fundamental question** in cooperative game theory.  
Another key problem in game theory is to try to understand what the outcomes of a game will be. An *outcome* of a game $G$ consists of two parts:  

1. A **coalition structure**, which is a partition of the gran coalition into coalitions $\big(CS = \{C_1,...,C_m\}\big)$
2. A **payoff vector** $\mathbb{x} = (x_1,...,x_p)$ associated to the coalition structure which distributes the value $\nu(C_j)$ of each coalition among its members. All its entries must be non negative and the sum of entries for players belonging to the same coalition must be lower or equal than the coalition value.

As one can imagine, **not all outcomes are equally desirable**. An outcome that pushes all agents together is (typically) preferable to an outcome that some agents want to deviate from and an outcome that shares the payoff in a fair way is desirable.  
There are two criteria to evaluate an outcome:

1. **Fairness**: how well each agent's payoff reflects its contribution.
2. **Stability**: what are the incentives for the agents to stay in the coalition structure.

We'll focus on the **Shapley Value** that belongs to a family of solution concepts which takes account mainly of the fairness of an outcome. It's defined with respect to the grand coalition: it defines a way of distributing the value $\nu(\mathcal{P})$ that could be obtained by the *gran coalition* (so, in the case where all players are in the same coalition), and it is based on the intuition that **the payment that each agent receives should be proportional to his contribution.**  
The initial idea is, given a **fixed ordering** of agents, to pay each one of them **according to how much is the increase in the value of the coalition formed by its precedessors**. The problem is that agents that play symmetric roles may receive different payoffs depending on their position in the ordering.  
The Shapley Value removes the dependence on the agents ordering by **averaging over all possible orderings (or permutations) of the players.**  
Some additional notation is needed to formally define the Shapley Value:

1. $S_\pi(j)$ is the set of all predecessors of player $j$ in the permutation $\pi \in \Pi_{\mathcal{P}}$ (where $\Pi_{\mathcal{P}}$ is the set of all permutations of $\mathcal{P}$). It's defined as $S_\pi(j)= \{r \in \mathcal{P} \ \text{such that} \ \pi(r) < \pi(j)\}$.  
2. **The marginal contribution** of an agent $j$ with respect to a permutation $\pi$ in a game $G = (\mathcal{P}, \nu)$ is $$\Delta_\pi^G(j) = \nu(S_\pi(j) \cup \{j\} ) - \nu(S_\pi(j))$$
and measures how much the player $j$ increases the value of the coalition consisting of its predecessor in $\pi$.  

Now, we can define the **Shapley Value** $\psi^G(j)$ (for a game $G$ and a player $j$) as **the average marginal contribution of the player $j$**, where the average is taken **over all permutations** of the player set (assumed to be **all equally likely**):

\begin{align}
\psi^G(j) = \mathbb{E}_{\pi}\left(\Delta_{\pi}^{G}(j)\right) = \frac{1}{p!} \sum_{\pi \in \Pi_{\mathcal{P}}} \Delta_{\pi}^{G}(j) = \sum_{C : j \notin C} \frac{(\#C)!(p-1-\#C)!}{p!}\left(\nu(C\cup\{j\})-\nu(C)\right)
\end{align}

Please note that the marginal contributions $\Delta_{\pi}^{G}(j)$ are of the form $\nu(C\cup\{j\})-\nu(C)$, where $C$ is a coalition not containing the player $j$. Also, there are $(\#C)!$ possible orderings of $C$ and $(p-1-\#C)!$ possible orderings of $\mathcal{P}\setminus(C\cup\{j\})$.  
It is shown that the Shapley Value is **the only payoff division scheme** that has all these four properties:

- **Efficiency**: $\sum_{j} \psi^G(j) = \nu(\mathcal{P})$.
- **Dummy player**: if player $j$ is dummy (i.e., $\nu(C \cup \{j\}) = \nu(C)$ for any coalition $C$), then $\psi^G(j) = 0$.
- **Symmetry**: if $j_1$ and $j_2$ are symmetric players (i.e., $\nu(C \cup \{j_1\}) = \nu(C \cup \{j_2\})$ for any coalition $C$), then $\psi^{G}(j_1) = \psi^{G}(j_2)$.  
- **Additivity**: given any two characteristic function games $G_1 = (\mathcal{P},\nu_1)$ and $G_2 = (\mathcal{P},\nu_2)$, and their sum $G^+ = (\mathcal{P},\nu_1 + \nu_2)$, $$\psi^+(j) = \psi^G_1(j) + \psi^G_2(j) \text{ for all } j \in \mathcal{P}.$$ 

In **simple games** (monotone and with characteristic function which only takes values 0 and 1), the Shapley Value measures the *power of the player*, the probability that it can influence the outcome of the game. To be more precise, it is the **probability that the player $j$ turns a losing coalition into a winning one.**  

The last important concept to stress is that the **Shapley value** can be used to determine the importance of the variables (**covariates**) involved in a supervised learning problem. It casts the problem as a game where the **covariates are players** and the **measure of fit is the characteristic function.** **The value of a coalition of covariates is the negative expected squared difference between the actual and predicted outcomes.** The Shapley value represents the **average contribution of a covariate to the predictive accuracy across all possible combinations of covariates.**   

An important remark regarding the Shapley value is that while it offers a comprehensive importance measure, **it doesn't come without limitations**, such as being affected by covariate correlation.

## Exercise 1: Get a better understanding of Games and Shapley

In this first exercise, given a game with 3 players, **Antwohnette (A), BadellPadel (B) and Chumbawamba (C)** we are asked to complete the definition of the **characteristic function game, $\nu$**, and then to calculate the **Shapley value**.  

Taking into account the definition of the **characteristic function game $\nu$** above we can claim that:  

- $\mathcal{P}=\{A,B,C\}$ is the set of our agents (our players).  
- $\nu: 2^\mathcal{P} \rightarrow \mathbb{R}$ is our **characteristic function** that we have to find.

From the text of the exercise we know that:

- $\nu(A)=\nu(B)=\nu(C)=0$    
- $\nu(ABC)=4$    

Now we have to find the other value of $\nu$ associated with the other combinations of our players, so we have to find $\nu(AB)$, $\nu(AC)$ and $\nu(BC)$.  
Let's start with the definition of our $\nu(C)$, that is **the numbers of hours potentially saved by well organized coalition**. In order to get our values we have to find the time window in which each pair of players saves time by working together and thus we must numerically count the hours in which our players are dedicated to solving the homework at the same time. 
From the setup of the game we know that:  

- **Player A works from 14:00 to 17:00**  
- **Player B works from 11:00 to 16:00**  
- **Player C works from 9:00 to 13:00** 

An so we can claim that:  

- $\nu(AB)=2$, because the hours in which both A and B are dedicated to solving the homework are exactly 2, from 14:00 to 16:00.   
- $\nu(AC)=0$, because A and C never have hours in common to work together on the homework and thus to 'team up'.  
- $\nu(BC)=2$, because the hours in which both A and B are dedicated to solving the homework are exactly 2, from 11:00 to 13:00.  

We have thus completed the definition of $\nu$ and we can proceed to calculate the **Shapley value**.  We have $\mathcal{P}=\{A,B,C\}$: the number of the permutations of a set of 3 elements is $3!=6$.  
Due to this we have to analyze the following $6$ permutations: $$\Pi_\mathcal{P}=\{(A,B,C), (A,C,B), (B,A,C), (B,C,A), (C,A,B), (C,B,A)\}$$ 
In order to get $\Delta_\pi^G(j) = \nu(S_\pi(j) \cup \{j\} ) - \nu(S_\pi(j))$ we have to analyze each permutation at time:

- $(A,B,C):$ 
    - $\Delta_\pi^G(A) = \nu(S_\pi(A) \cup \{A\} ) - \nu(S_\pi(A))= \nu(\emptyset \cup \{A\} ) - \nu(\emptyset) = \nu(A) - \nu(\emptyset)=0$   
    - $\Delta_\pi^G(B) = \nu(S_\pi(B) \cup \{B\} ) - \nu(S_\pi(B))=\nu(\{A\} \cup \{B\} ) - \nu(A) = \nu(AB) - \nu(A)=2-0=2$  
    - $\Delta_\pi^G(C) = \nu(S_\pi(C) \cup \{C\} ) - \nu(S_\pi(C))= \nu(\{A,B\} \cup \{C\} ) - \nu(AB) = \nu(ABC) - \nu(AB)=4-2=2$  
- $(A,C,B):$
    - $\Delta_\pi^G(A) = 0.$  
    Because the calculation here is the same of the the previous step because the $A$ is in the same position in the permutation so its **marginal contribution** is the same.
    - $\Delta_\pi^G(C) = \nu(S_\pi(C) \cup \{C\} ) - \nu(S_\pi(C))= \nu(\{A\} \cup \{C\} ) - \nu(A) = \nu(AC) - \nu(A)=0-0=0$  
    - $\Delta_\pi^G(B) = 4$  
    We can claim this because $\sum_{j\in \pi}\Delta_\pi^G(j)=\nu(A,B,C)=4$ so since we know that $\Delta_\pi^G(A) = 0$ and $\Delta_\pi^G(C) =0$ definitely $\Delta_\pi^G(B) = 4.$  
- $(B,A,C):$
    - $\Delta_\pi^G(B) = \nu(S_\pi(B) \cup \{B\} ) - \nu(S_\pi(B))=\nu(\emptyset \cup \{B\} ) - \nu(\emptyset) = \nu(B) - \nu(\emptyset)=0-0=0$  
    - $\Delta_\pi^G(A) = \nu(S_\pi(A) \cup \{A\} ) - \nu(S_\pi(A))=\nu(\{B\} \cup \{A\} ) - \nu(B)= \nu(AB) - \nu(B)=2-0=2$  
    - $\Delta_\pi^G(C) = 2$  
    The reason behind this result is the same of the previous step $\big(\sum_{j\in \pi}\Delta_\pi^G(j)=\nu(A,B,C)=4\big).$    
- $(B,C,A):$
    - $\Delta_\pi^G(B) = 0$  
    Because the calculation here is the same of the the previous step because the $B$ is in the same position in the permutation so its **marginal contribution** is the same.
    - $\Delta_\pi^G(C) = \nu(S_\pi(C) \cup \{C\} ) - \nu(S_\pi(C)) = \nu(\{B\} \cup \{C\} ) - \nu(B)=\nu(BC) - \nu(B) = 2-0=2$  
    - $\Delta_\pi^G(A) = 2$  
    Due to the fact that $\sum_{j\in \pi}\Delta_\pi^G(j)=\nu(A,B,C)=4$ we have this result.  
- $(C,A,B):$  
    - $\Delta_\pi^G(C) = \nu(S_\pi(C) \cup \{C\} ) - \nu(S_\pi(C))=\nu(\emptyset \cup \{C\} ) - \nu(\emptyset) = \nu(C) - \nu(\emptyset)=0-0=0$ 
    - $\Delta_\pi^G(A) = \nu(S_\pi(A) \cup \{A\} ) - \nu(S_\pi(A))=\nu(\{C\} \cup \{A\} ) - \nu(C) = \nu(AC) - \nu(C)=0-0=0$ 
    - $\Delta_\pi^G(B) = 4$  
    Here again the result it is this due to the fact that $\sum_{j\in \pi}\Delta_\pi^G(j)=\nu(A,B,C)=4.$
- $(C,B,A):$  
    - $\Delta_\pi^G(C) = 0$  
    Because the calculation here is the same of the the previous step because the $C$ is in the same position in the permutation so its **marginal contribution** is the same. 
    - $\Delta_\pi^G(B) = \nu(S_\pi(B) \cup \{B\} ) - \nu(S_\pi(B))=\nu(\{C\} \cup \{B\} ) - \nu(C) = \nu(BC) - \nu(C)=2-0=2$
    - $\Delta_\pi^G(A) = 2$  
    Here again the result it is this due to the fact that $\sum_{j\in \pi}\Delta_\pi^G(j)=\nu(A,B,C)=4.$  
    
So to summarize all the calculations in a table we have:  

\[
\begin{array}{c|ccc}
\text{Permutation} & \text{Player A} & \text{Player B} & \text{Player C} \\ \hline
ABC & 0 & 2 & 2 \\
ACB & 0 & 4 & 0 \\
BAC & 2 & 0 & 2 \\
BCA & 2 & 0 & 2 \\
CAB & 0 & 4 & 0 \\
CBA & 2 & 2 & 0 \\ \hline
\text{TOTAL VALUE} & 6 & 12 & 6 \\
\end{array}
\]

At this point, we have everything to calculate the **Shapley value**. First of all, remember that the **Shapley value** of each player $j$ (where $j \in \{A,B,C\}$ in our case) is the function that calculates the average marginal contribution of each players, taken over all permutations of the player-set, in particular we have: $$\psi^G(j)=\mathbb{E}[\Delta_\pi^G(j)]=\frac{1}{p!}\sum_{\pi \in \Pi_\mathcal{P}} \Delta_\pi^G(j)$$ 
In our scenario we have $3!=6$ permutations and for each player we have:  
- **Player A:** $\sum_{\pi \in \Pi_\mathcal{P}} \Delta_\pi^G(A)=6$  
- **Player B:** $\sum_{\pi \in \Pi_\mathcal{P}} \Delta_\pi^G(B)=12$    
- **Player C:** $\sum_{\pi \in \Pi_\mathcal{P}} \Delta_\pi^G(C)=6$  
So in the end we can claim that:  
$$\psi^G(A)=\frac{6}{6}=1 \\ \psi^G(B)=\frac{12}{6}=2 \\ \psi^G(C)=\frac{6}{6}=1$$ 
**Interpreting this value in the given context suggests us that Antwohnette and Chumbawamba can potentially save 1 hour by forming a well organized coalition, while BadellPadel can save 2 hours.**

## Exercise 2: WIP
In this exercise it is asked to leverage the theory above in a financial context. There is a **portfolio of $p$ stocks** whose returns are modeled by a set of random variables $\{X_1,...,X_p\}$. The **utility function** is defined as follows:
\begin{align}
U_{\omega}(X) = \mathbb{E}(X) - \omega\cdot\mathbb{V}ar(X) \quad \text{for some weight} \quad \omega > 0
\end{align}
It's basically a linear combination of the average return and volatility of the portfolio. The problem consists in deciding how to allocate the capital among all of the stocks: intuitively, it would be good that the allocation is proportional to the asset's contribution to the total utility $U_{\omega}(\sum_{j=1}^pX_j)$.  
We now define this setting in a **cooperative game theory prespective**. Let the assets be the players and let $C$ be any coalition of them, with $R_C = \sum_{j\in C}X_j$ being their return. For any $C\subset\mathcal{P}$ we define the characteristic function as follows:
\begin{align}
\mathcal{v}(C) = U_{\omega}(R_C) = \mathbb{E}(R_C) - \omega\cdot\mathbb{V}ar(R_C) = \mathcal{v}_{\mathbb{E}}(C) - \omega\cdot\mathcal{v}_{\mathbb{V}}(C)
\end{align}
Hence, the game we're defining $(G = (\mathcal{P},\mathcal{v}))$ is a linear combination of two other games: $G_{\mathbb{E}} = (\mathcal{P},\mathcal{v}_{\mathbb{E}})$ and $G_{\mathbb{V}} = (\mathcal{P},\mathcal{v}_{\mathbb{V}})$.  
Since our goal is to allocate the capital proportionally to the contribution given to the total utility, the Shapley Value turns out to be suitable and simple.  
By the **additivity of the Shapley Value** we get that:
\begin{align}
\psi^G(j) = \psi^G_{\mathbb{E}}(j) - \omega\cdot \psi^G_{\mathbb{V}}(j)
\end{align}
By the **linearity of expectation**, the first term is:
\begin{align}
\psi^G_{\mathbb{E}}(j) = \mathbb{E}(X_j)
\end{align}
While the second term, after elementary manipulations, turns out to be:
\begin{align}
\psi^G_{\mathbb{V}}(j) = \mathbb{C}ov(X_j,R_{\mathcal{P}})
\end{align}
In conclusion, after picking a suitable weight $\omega > 0$, the Shapley allocation to each stock is given by:
\begin{align}
\psi^G(j) = \mathbb{E}(X_j) - \omega\cdot\mathbb{C}ov(X_j,R_{\mathcal{P}}) = \mathbb{E}(X_j) - \omega\cdot\sum_{r=1}^p\mathbb{C}ov(X_j,X_r)
\end{align}
In other words, the **Shapley Value used to allocate capital is a population quantity to be estimated**. By taking a look at the formula, it can be notices that we would need to study the marginal correlation graph among some standard measure of stock *relative performance* (which may be calculated starting from the *daily closing prices* of the stocks).  
In practice, we need to:  

- Select a sensibly sized portfolio of $p$ stocks and take their *closing price* from **January 1, 2020 till the end of 2023.**
- From the closing prices $c_{t,j}$ (stock $j$ on day $t$), we have to **compute their relative variation** WIP $x_{t,j} = \log(c_{t,j}/c_{t-1,j})$.
- By seeing each stock as a node, we have to **estimate the Pearson-based correlation graph over stocks**, by treating the instances $\{x_{t,j}\}_t$ as independent samples (even though they're not). We place an undirected edge between stocks $j1$ and $j2$ only if their estimated Pearson correlation is large enough (statistically significative or over a threshold $\tau > 0$). Beware of multiplicity when choosing the threshold! 
- After **properly visualizing the graph**, we have to draw some conclusions based on the decisions made along the way.
- Finally, we have to **provide and visualize bootstrapped confidence intervals for the Shapley Values** of the stocks in our portfolio and comment the results based on the choices we've made.

Energy, Information Technology, health care, Consumer Discrectionary, Financials. Due per settore

- ENI (ITALY)
- SUMISEKI HOLDINGS (JAPAN)
- APPLE (USA)
- MICROSOFT (USA)
- MC DONALDS (USA)
- CARREFOUR (FRANCIA)
- BAYER (GERMANIA)
- CIPLA (INDIA)
- BLACKROCK (USA)
- JP MORGAN (USA)



WIP - DESCRIBE THE STOCKS CHOSEN AND THE LABELS ASSIGNED
```{r, echo=TRUE}
# PREPROCESSING PART 1
# Get stocks data with the library 'tseries'
suppressPackageStartupMessages(library(tseries, quietly=T))

# Let's pull data about the chosen stocks
eni <- get.hist.quote(instrument="E", start="2020-01-01", end="2023-12-31", quote= c("Open","Close"), provider="yahoo", drop=TRUE, quiet = TRUE)

smh <- get.hist.quote(instrument="1514.T", start="2020-01-01", end="2023-12-31", quote= c("Open","Close"), provider="yahoo", drop=TRUE, quiet = TRUE)

aapl <- get.hist.quote(instrument="AAPL", start="2020-01-01", end="2023-12-31", quote= c("Open","Close"), provider="yahoo", drop=TRUE, quiet = TRUE)

micro <- get.hist.quote(instrument="MSFT", start="2020-01-01", end="2023-12-31", quote= c("Open","Close"), provider="yahoo", drop=TRUE, quiet = TRUE)

mcd <- get.hist.quote(instrument="MCD", start="2020-01-01", end="2023-12-31", quote= c("Open","Close"), provider="yahoo", drop=TRUE, quiet = TRUE)

carf <- get.hist.quote(instrument="CA.PA", start="2020-01-01", end="2023-12-31", quote= c("Open","Close"), provider="yahoo", drop=TRUE, quiet = TRUE)

bay <- get.hist.quote(instrument="BAYN.DE", start="2020-01-01", end="2023-12-31", quote= c("Open","Close"), provider="yahoo", drop=TRUE, quiet = TRUE)

cipla <- get.hist.quote(instrument="CIPLA.NS", start="2020-01-01", end="2023-12-31", quote= c("Open","Close"), provider="yahoo", drop=TRUE, quiet = TRUE)

brk <- get.hist.quote(instrument="BLK", start="2020-01-01", end="2023-12-31", quote= c("Open","Close"), provider="yahoo", drop=TRUE, quiet = TRUE)

jpm <- get.hist.quote(instrument="JPM", start="2020-01-01", end="2023-12-31", quote= c("Open","Close"), provider="yahoo", drop=TRUE, quiet = TRUE)

# Fill eventual holidays to prevent different series' length
eni <- zoo::na.approx(eni, xout = seq(start(smh), end(eni), by = "day"), method = "constant")
smh <- zoo::na.approx(smh, xout = seq(start(smh), end(smh), by = "day"), method = "constant")
aapl <- zoo::na.approx(aapl, xout = seq(start(smh), end(aapl), by = "day"), method = "constant")
micro <- zoo::na.approx(micro, xout = seq(start(smh), end(micro), by = "day"), method = "constant")
mcd <- zoo::na.approx(mcd, xout = seq(start(smh), end(mcd), by = "day"), method = "constant")
carf <- zoo::na.approx(carf, xout = seq(start(smh), end(carf), by = "day"), method = "constant")
bay <- zoo::na.approx(bay, xout = seq(start(smh), end(bay), by = "day"), method = "constant")
cipla <- zoo::na.approx(cipla, xout = seq(start(smh), end(cipla), by = "day"), method = "constant")
brk <- zoo::na.approx(brk, xout = seq(start(smh), end(brk), by = "day"), method = "constant")
jpm <- zoo::na.approx(jpm, xout = seq(start(smh), end(jpm), by = "day"), method = "constant")

# Compute their relative log-returns
eni$y = diff(log(eni$Close))
smh$y = diff(log(smh$Close))
aapl$y = diff(log(aapl$Close))
micro$y = diff(log(micro$Close))
mcd$y = diff(log(mcd$Close))
carf$y = diff(log(carf$Close))
bay$y = diff(log(bay$Close))
cipla$y = diff(log(cipla$Close))
brk$y = diff(log(brk$Close))
jpm$y = diff(log(jpm$Close))
```

```{r, echo=FALSE, fig.width=10, fig.height=8, fig.align='center', out.width = '200%', out.height = '400%', dpi = 300}
# Now, let's visualize the closing prices of the chosen stocks along with their relative log-returns
plot(cbind(eni$Close,smh$Close,aapl$Close,micro$Close,mcd$Close,carf$Close,bay$Close,cipla$Close,brk$Close,jpm$Close), main = 'Stocks closing prices', xlab = 'Date', ylab = c('ENI','SMH','AAPL','MICRO','MCD','CARF','BAY','CIPLA','BRK','JPM'), cex.main = 2)
plot(cbind(eni$y,smh$y,aapl$y,micro$y,mcd$y,carf$y,bay$y,cipla$y,brk$y,jpm$y), main = 'Stocks relative log-returns', xlab = 'Date', ylab = c('ENI','SMH','AAPL','MICRO','MCD','CARF','BAY','CIPLA','BRK','JPM'), cex.main = 2)
```

```{r, echo=TRUE}
# PREPROCESSING PART 2
# Create a dataframe of the relative log-returns for each stock
# We've decided to drop the first value as the relative log-return can't be calculated for the first day
eni <- as.data.frame(eni$y[-1])
eni$dates <- rownames(eni)
smh <- as.data.frame(smh$y[-1])
smh$dates <- rownames(smh)
aapl <- as.data.frame(aapl$y[-1])
aapl$dates <- rownames(aapl)
micro <- as.data.frame(micro$y[-1])
micro$dates <- rownames(micro)
mcd <- as.data.frame(mcd$y[-1])
mcd$dates <- rownames(mcd)
carf <- as.data.frame(carf$y[-1])
carf$dates <- rownames(carf)
bay <- as.data.frame(bay$y[-1])
bay$dates <- rownames(bay)
cipla <- as.data.frame(cipla$y[-1])
cipla$dates <- rownames(cipla)
brk <- as.data.frame(brk$y[-1])
brk$dates <- rownames(brk)
jpm <- as.data.frame(jpm$y[-1])
jpm$dates <- rownames(jpm)

# Join the datasets together to get a data matrix of their relative log-returns
df <- plyr::join_all(list(eni,smh,aapl,micro,mcd,carf,bay,cipla,brk,jpm), by = 'dates', type = 'full')
df <- subset(df, select = -c(dates))
colnames(df) <- c('eni','smh','aapl','micro','mcd','carf','bay','cipla','brk','jpm')
```

Now we have to build the marginal correlation graph (thus to choose the threshold)... WIP
```{r, echo=TRUE}
# CHOOSING THE THRESHOLD
# Calculate the correlation matrix between the relative log-returns of the stocks
corr_mat <- cor(df)

# We will add an edge connecting two stocks if the associated Pearson's correlation Confidence Interval doesn't intersect a symmetric interval [-tau,tau]. We've decided that tau is the lowest statistically significant Pearson's correlation coefficient overall. Since we need to understand the *global* topology of the graph, the statistical significance must be asserted by correcting the level with Bonferroni method.

# In our case there are m=45 different Pearson's correlation coefficients
# Hence, starting from a level alpha=0.05, the (Bonferroni) corrected level is
m = 45
alpha = 0.05
alpha_corrected <- alpha/m # 0.001

# Get a matrix with p-values for the calculated Pearson's correlations
cor.test.p <- function(x){
    PVAL <- function(x, y) cor.test(x, y)[['p.value']]
    pvals <- outer(
      colnames(x), 
      colnames(x), 
      Vectorize(function(i,j) PVAL(x[,i], x[,j]))
    )
    dimnames(pvals) <- list(colnames(x), colnames(x))
    return(pvals)
}

# Get a matrix containing the lower bounds of the Confidence Intervals for Pearson's correlation (at the corrected level) 
cor.test.lower <- function(x){
    LOW <- function(x, y) cor.test(x, y, conf.level = 1-alpha_corrected)[['conf.int']][1]
    lower_bounds <- outer(
      colnames(x), 
      colnames(x), 
      Vectorize(function(i,j) LOW(x[,i], x[,j]))
    )
    dimnames(lower_bounds) <- list(colnames(x), colnames(x))
    return(lower_bounds)
}

# Get a matrix containing the upper bounds of the Confidence Intervals for Pearson's correlation (at the corrected level)
cor.test.upper <- function(x){
    UP <- function(x, y) cor.test(x, y, conf.level = 1-alpha_corrected)[['conf.int']][2]
    upper_bounds <- outer(
      colnames(x), 
      colnames(x), 
      Vectorize(function(i,j) UP(x[,i], x[,j]))
    )
    dimnames(upper_bounds) <- list(colnames(x), colnames(x))
    return(upper_bounds)
}

pvals <- cor.test.p(df)
lower_b <- cor.test.lower(df)
upper_b <- cor.test.upper(df)

# The threshold is the minimum correlation coefficient among the statistically significant ones
tau <- min(corr_mat[which(pvals < alpha_corrected)])

# The edges which are going to be added to the graph are those corresponding to a Pearson's correlation coefficient with a CI non intersecting the interval [-tau,tau]. By doing this selection maintaining the structure of the correlation matrix we build an adjacency matrix for the graph
edges <- (upper_b < -tau | lower_b > tau)
# Build the adjacency matrix with correlations as weights
adj_mat <- matrix(ifelse(edges, round(corr_mat,2), 0), nrow = 10, ncol = 10)

#LA THRESHOLD È LA CORRELATION SIGNIFICATIVA PIÙ BASSA DI TUTTE - WIP

# WIP - Prettify it
knitr::kable(corr_mat ,  caption = "Correlation Matrix")
```

Here is the graph obtained WIP
```{r, echo=FALSE, fig.width=10, fig.height=8, fig.align='center', out.width = '200%', out.height = '400%', dpi = 300}
# GRAPH VISUALIZATION
# Import igraph library and use it to create the graph, which is further simplified to remove duplicates and self loops
suppressPackageStartupMessages(library(igraph, quietly=T))
g_corr <- simplify(graph_from_adjacency_matrix(adj_mat, mode = "undirected", weighted = T))

# Add node labels to make the visualization prettier
V(g_corr)$label <- c('ENI','SUMISEKI','APPLE','MICROSOFT','MC DONALDS','CARREFOUR','BAYER','CIPLA','BLACKROCK','JP MORGAN')
# Create a color map for the countries
V(g_corr)$country <- c('Italy','Japan','USA','USA','USA','France','Germany','India','USA','USA')
col <- data.frame(country = unique(V(g_corr)$country), stringsAsFactors = F)
col$color <- c('darkgreen','white','red','royalblue','black','orange')
# Attach the colors to the nodes
V(g_corr)$color <- col$color[match(V(g_corr)$country, col$country)]

# It's crucial to set this graphical parameter to avoid giant blank spaces at the margin of the graph
par(mar=c(1.5,0,4.5,0)+.1)

# Visualize the graph with an appropriate layout
# We've chosen to layout them as a star, placing at the center the node with the highest strength (sum of its edges' weights)
h_str_id <- V(g_corr)[which(strength(g_corr) == sort(strength(g_corr), decreasing = TRUE)[1])]
layoutcorr <- layout_as_star(g_corr, center = h_str_id)
layoutcorr <- norm_coords(layoutcorr, ymin=-1, ymax=1, xmin=-1, xmax=1)

plot(g_corr,
     rescale = F,
     layout = layoutcorr, # Personalized layout
     main = "Pearson's correlation Graph",
     edge.arrow.size = 1, # Arrow size
     edge.curved = 0.2, # Arcs curvature
     edge.color = 'black',
     edge.width = 1,
     vertex.size = 5 + strength(g_corr)*3, # Nodes size (proportional to their strength)
     vertex.frame.color = 'black',
     vertex.frame.width = 2,
     vertex.label = V(g_corr)$label, # Decide whether to use the attribute "label" from nodes
     vertex.label.cex = 0.7 + strength(g_corr)/10, # Nodes' labels size
     vertex.label.color='black',
     vertex.label.dist=2.1,
     vertex.label.degree = c(0,-(2/9)*pi,-(4/9)*pi,-(2/3)*pi,-(7/9)*pi,-(11/9)*pi,-(13/9)*pi,-(15/9)*pi,-pi/2,-(15/9)*pi),
     vertex.label.font = 2) # Bold font 

legend("topright", legend = col$country, pt.bg=col$color, bty="n",
       pch=21, col="#777777", cex = 1.2, title = as.expression(bquote(bold("Country"))), title.adj = 0.5)

# Reset the graphical parameter to default
par(mar=c(5, 4, 4, 2) + 0.1)
```

It is clear that the relative log-returns of the stocks are strongly(? WIP) correlated when they belong to the same country: the american stocks in our case are an example. It can be seen also that the relative log-returns of european stocks are correlated among them and with some american stocks, suggesting the hypothesis that the american market can partly influence the european one. The relative log-returns of the only two stocks from asian countries are correlated with no other stock, highlighting the degree of independence of their markets.

### Bootstrap CIs for the Shapley Values
WIP bla bla bla introduction
```{r, echo=TRUE}
# Bootstrap CIs for Shapley Values
B = 10000 # Number of iterations for bootstrap
n <- nrow(df)
shap.B <- matrix(NA, nrow = B, ncol = ncol(df)) # Bootstrap B shapley values for each stock
dimnames(shap.B) <- list(1:B, colnames(df)) # Rename columns
omega <- 0.5 # Select the risk propensity
set.seed(4545) # Set seed for reproducibility
# Main loop - non-parametric bootstrap for Shapley Values
for (b in 1:B){
  idx <-  sample(1:n, replace = T)
  cov_mat <- matrix(ifelse(edges, cov(df[idx,]), 0), nrow = 10, ncol = 10) # Create a covariance matrix, fixing the covariance to zero for variables not linked in the correlation graph - WIP
  for (j in 1:10){
    shap.B[b,j] <- mean(df[idx,j]) - omega*sum(cov_mat[j,]) # shapley formula, we use the sample mean as estimator of the expected value WIP and the adjacency matrix of the correlation graph to retrieve the covariance between the variables
  }
}

# Calculate percentile CIs for the Shapley Values and store them in a matrix
perc_int <- function(bootsamp, alpha){
  lows <- c(rep(NA,ncol(bootsamp)))
  ups <- c(rep(NA,ncol(bootsamp)))
  for (j in 1:ncol(bootsamp)){
    lows[j] <- unname(quantile(bootsamp[,j], alpha/2))
    ups[j] <-  unname(quantile(bootsamp[,j], 1-alpha/2))
  }
  out <- matrix(c(lows,ups), ncol = 2)
  dimnames(out) <- list(colnames(bootsamp), c('Lower Bound','Upper Bound'))
  return(out)
}

shap_ci_perc <- perc_int(shap.B,0.05)

# WIP - Prettify it
knitr::kable(shap_ci_perc ,  caption = "Bootstrapped Confidence Intervals for the Shapley Value of each stock")
```

WIP - Write comment, test several values of omega and select alpha wisely. Also, check if the mean is ok to estimate the expected value of the log-returns