---
title: "Homework 3 - Group 46"
author: "Name and Surname: Alessandra Colaiocco - ID: 1843714 - Email: colaiocco.1843714@studenti.uniroma1.it \n\n Name and Surname: Gabriele Parisini - ID: 1964381 - Email: parisini.1964381@studenti.uniroma1.it"
output: html_document
date: "2024-01-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Game Theory and Shapley Value
## Introduction
A *game* (in *game theory*) is an abstract mathematical model of a scenario in which *agents* (or *players*) interact. When we're dealing with a game, we're interested in those features of the scenario that are relevant to the decisions of the players.  
In this homework we're focusing on **cooperative games**, which are games populated by a set $\mathcal{P} = \{1,...,p\}$ of players (or *grand coalition*) which are able to form *coalitions* (subsets of the "grand" one).  
To continue further in the explanation, we would need to define the *characteristic function game*:  
  
**Definition 1.**&nbsp;&nbsp;A characteristic function game $G$ is given by a pair $(\mathcal{P},\mathcal{v})$, where $\mathcal{P} = \{1,...,p\}$ is a non-empty set and $\mathcal{v}: 2^{\mathcal{P}}\mapsto\mathbb{R}$ is a **characteristic function** which maps every coalition $C \subseteq\mathcal{P}$ to a real number $\mathcal{v}(C)$ (which is referred to as the **value of the coalition $C$**).  
  
It's crucial to understand that the characteristic function game assigns a value to a coalition as a whole, **not telling how the coalition value should be divided among the members of the coalition itself.** How to divide the coalition value is a **fundamental question** in cooperative game theory.  
Another key problem in game theory is to try to understand what the outcomes of a game will be. An *outcome* of a game $G$ consists of two parts:  

1. A **coalition structure**, which is a partition of the gran coalition into coalitions ($CS = \{C_1,...,C_m\}$)
2. A **payoff vector** $\mathbb{x} = (x_1,...,x_p)$ associated to the coalition structure which distributes the value $\mathcal{v}(C_j)$ of each coalition among its members. All its entries must be non negative and the sum of entries for players belonging to the same coalition must be lower or equal than the coalition value.

As one can imagine, **not all outcomes are equally desirable**. An outcome that pushes all agents together is (typically) preferable to an outcome that some agents want to deviate from and an outcome that shares the payoff in a fair way is desirable.  
There are two criteria to evaluate an outcome:

1. **Fairness**: how well each agent's payoff reflects its contribution.
2. **Stability**: what are the incentives for the agents to stay in the coalition structure.

The **Shapley Value** belongs to a family of solution concepts which takes account mainly of the fairness of an outcome. It's defined with respect to the grand coalition: it defines a way of distributing the value $\mathcal{v}(\mathcal{P})$ that could be obtained by the *gran coalition* (so, in the case where all players are in the same coalition), and it is based on the intuition that **the payment that each agent receives should be proportional to his contribution.**  
The initial idea is, given a **fixed ordering** of agents, to pay each one of them **according to how much is the increase in the value of the coalition formed by its precedessors**. The problem is that agents that play symmetric roles may receive different payoffs depending on their position in the ordering.  
The Shapley Value removes the dependence on the agents ordering by **averaging over all possible orderings (or permutations) of the players.**  
Some additional notation is needed to formally define the Shapley Value:

1. WIP
2. WIP

Now, we can define the Shapley Value $\psi^G(j)$ (for a game $G$ and a player $j$) as **the average marginal contribution of the player $j$**, where the average is taken **over all permutations** of the player set (assumed to be **all equally likely**):

\begin{align}
\psi^G(j) = \mathbb{E}_{\pi}\left(\Delta_{\pi}^{G}(j)\right) = \frac{1}{p!} \sum_{\pi \in \Pi_{\mathcal{P}}} \Delta_{\pi}^{G}(j) = \sum_{C : j \notin C} \frac{(\#C)!(p-1-\#C)!}{p!}\left(\mathcal{v}(C\cup\{j\})-\mathcal{v}(C)\right)
\end{align}

Please note that the marginal contributions $\Delta_{\pi}^{G}(j)$ are of the form $\mathcal{v}(C\cup\{j\})-\mathcal{v}(C)$, where $C$ is a coalition not containing the player $j$. Also, there are $(\#C)!$ possible orderings of $C$ and $(p-1-\#C)!$ possible orderings of $\mathcal{P}\setminus(C\cup\{j\})$.  
It is shown that the Shapley Value is **the only payoff division scheme** that has all these four properties:

- **Efficiency**: WIP
- **Dummy player**: WIP
- **Symmetry**: WIP 
- **Additivity**: WIP

In **simple games** (monotone and with characteristic function which only takes values 0 and 1), the Shapley Value measures the *power of the player*, the probability that it can influence the outcome of the game. To be more precise, it is the **probability that the player $j$ turns a losing coalition into a winning one.**  
WIP - BRIEF DESCRIPTION OF SHAPLEY VALUE FOR VARIABLE IMPORTANCE  

## Exercise 1: Get a better understanding of Games and Shapley
WIP

## Exercise 2: WIP
In this exercise it is asked to leverage the theory above in a financial context. There is a **portfolio of $p$ stocks** whose returns are modeled by a set of random variables $\{X_1,...,X_p\}$. The **utility function** is defined as follows:
\begin{align}
U_{\omega}(X) = \mathbb{E}(X) - \omega\cdot\mathbb{V}ar(X) \quad \text{for some weight} \quad \omega > 0
\end{align}
It's basically a linear combination of the average return and volatility of the portfolio. The problem consists in deciding how to allocate the capital among all of the stocks: intuitively, it would be good that the allocation is proportional to the asset's contribution to the total utility $U_{\omega}(\sum_{j=1}^pX_j)$.  
We now define this setting in a **cooperative game theory prespective**. Let the assets be the players and let $C$ be any coalition of them, with $R_C = \sum_{j\in C}X_j$ being their return. For any $C\subset\mathcal{P}$ we define the characteristic function as follows:
\begin{align}
\mathcal{v}(C) = U_{\omega}(R_C) = \mathbb{E}(R_C) - \omega\cdot\mathbb{V}ar(R_C) = \mathcal{v}_{\mathbb{E}}(C) - \omega\cdot\mathcal{v}_{\mathbb{V}}(C)
\end{align}
Hence, the game we're defining $(G = (\mathcal{P},\mathcal{v}))$ is a linear combination of two other games: $G_{\mathbb{E}} = (\mathcal{P},\mathcal{v}_{\mathbb{E}})$ and $G_{\mathbb{V}} = (\mathcal{P},\mathcal{v}_{\mathbb{V}})$.  
Since our goal is to allocate the capital proportionally to the contribution given to the total utility, the Shapley Value turns out to be suitable and simple.  
By the **additivity of the Shapley Value** we get that:
\begin{align}
\psi^G(j) = \psi^G_{\mathbb{E}}(j) - \omega\cdot \psi^G_{\mathbb{V}}(j)
\end{align}
By the **linearity of expectation**, the first term is:
\begin{align}
\psi^G_{\mathbb{E}}(j) = \mathbb{E}(X_j)
\end{align}
While the second term, after elementary manipulations, turns out to be:
\begin{align}
\psi^G_{\mathbb{V}}(j) = \mathbb{C}ov(X_j,R_{\mathcal{P}})
\end{align}
In conclusion, after picking a suitable weight $\omega > 0$, the Shapley allocation to each stock is given by:
\begin{align}
\psi^G(j) = \mathbb{E}(X_j) - \omega\cdot\mathbb{C}ov(X_j,R_{\mathcal{P}}) = \mathbb{E}(X_j) - \omega\cdot\sum_{r=1}^p\mathbb{C}ov(X_j,X_r)
\end{align}
In other words, the **Shapley Value used to allocate capital is a population quantity to be estimated**. By taking a look at the formula, it can be notices that we would need to study the marginal correlation graph among some standard measure of stock *relative performance* (which may be calculated starting from the *daily closing prices* of the stocks).  
In practice, we need to:  

- Select a sensibly sized portfolio of $p$ stocks and take their *closing price* from **January 1, 2020 till the end of 2023.**
- From the closing prices $c_{t,j}$ (stock $j$ on day $t$), we have to **compute their relative variation** WIP $x_{t,j} = \log(c_{t,j}/c_{t-1,j})$.
- By seeing each stock as a node, we have to **estimate the Pearson-based correlation graph over stocks**, by treating the instances $\{x_{t,j}\}_t$ as independent samples (even though they're not). We place an undirected edge between stocks $j1$ and $j2$ only if their estimated Pearson correlation is large enough (statistically significative or over a threshold $\tau > 0$). Beware of multiplicity when choosing the threshold! 
- After **properly visualizing the graph**, we have to draw some conclusions based on the decisions made along the way.
- Finally, we have to **provide and visualize bootstrapped confidence intervals for the Shapley Values** of the stocks in our portfolio and comment the results based on the choices we've made.

Energy, Information Technology, health care, Consumer Discrectionary, Financials. Due per settore

- ENI (ITALY)
- SUMISEKI HOLDINGS (JAPAN)
- APPLE (USA)
- MICROSOFT (USA)
- MC DONALDS (USA)
- CARREFOUR (FRANCIA)
- BAYER (GERMANIA)
- CIPLA (INDIA)
- BLACKROCK (USA)
- JP MORGAN (USA)



WIP - DESCRIBE THE STOCKS CHOSEN AND THE LABELS ASSIGNED
```{r, echo=TRUE}
# PREPROCESSING PART 1
# Get stocks data with the library 'tseries'
library(tseries, quietly = T)
options("getSymbols.warning4.0" = FALSE)

# Let's pull data about the chosen stocks
eni <- get.hist.quote(instrument="E", start="2020-01-01", end="2023-12-31", quote= c("Open","Close"), provider="yahoo", drop=TRUE, quiet = TRUE)

smh <- get.hist.quote(instrument="1514.T", start="2020-01-01", end="2023-12-31", quote= c("Open","Close"), provider="yahoo", drop=TRUE, quiet = TRUE)

aapl <- get.hist.quote(instrument="AAPL", start="2020-01-01", end="2023-12-31", quote= c("Open","Close"), provider="yahoo", drop=TRUE, quiet = TRUE)

micro <- get.hist.quote(instrument="MSFT", start="2020-01-01", end="2023-12-31", quote= c("Open","Close"), provider="yahoo", drop=TRUE, quiet = TRUE)

mcd <- get.hist.quote(instrument="MCD", start="2020-01-01", end="2023-12-31", quote= c("Open","Close"), provider="yahoo", drop=TRUE, quiet = TRUE)

carf <- get.hist.quote(instrument="CA.PA", start="2020-01-01", end="2023-12-31", quote= c("Open","Close"), provider="yahoo", drop=TRUE, quiet = TRUE)

bay <- get.hist.quote(instrument="BAYN.DE", start="2020-01-01", end="2023-12-31", quote= c("Open","Close"), provider="yahoo", drop=TRUE, quiet = TRUE)

cipla <- get.hist.quote(instrument="CIPLA.NS", start="2020-01-01", end="2023-12-31", quote= c("Open","Close"), provider="yahoo", drop=TRUE, quiet = TRUE)

brk <- get.hist.quote(instrument="BLK", start="2020-01-01", end="2023-12-31", quote= c("Open","Close"), provider="yahoo", drop=TRUE, quiet = TRUE)

jpm <- get.hist.quote(instrument="JPM", start="2020-01-01", end="2023-12-31", quote= c("Open","Close"), provider="yahoo", drop=TRUE, quiet = TRUE)

# Fill eventual holidays to prevent different series' length
eni <- zoo::na.approx(eni, xout = seq(start(smh), end(eni), by = "day"), method = "constant")
smh <- zoo::na.approx(smh, xout = seq(start(smh), end(smh), by = "day"), method = "constant")
aapl <- zoo::na.approx(aapl, xout = seq(start(smh), end(aapl), by = "day"), method = "constant")
micro <- zoo::na.approx(micro, xout = seq(start(smh), end(micro), by = "day"), method = "constant")
mcd <- zoo::na.approx(mcd, xout = seq(start(smh), end(mcd), by = "day"), method = "constant")
carf <- zoo::na.approx(carf, xout = seq(start(smh), end(carf), by = "day"), method = "constant")
bay <- zoo::na.approx(bay, xout = seq(start(smh), end(bay), by = "day"), method = "constant")
cipla <- zoo::na.approx(cipla, xout = seq(start(smh), end(cipla), by = "day"), method = "constant")
brk <- zoo::na.approx(brk, xout = seq(start(smh), end(brk), by = "day"), method = "constant")
jpm <- zoo::na.approx(jpm, xout = seq(start(smh), end(jpm), by = "day"), method = "constant")

# Compute their relative log-returns
eni$y = diff(log(eni$Close))
smh$y = diff(log(smh$Close))
aapl$y = diff(log(aapl$Close))
micro$y = diff(log(micro$Close))
mcd$y = diff(log(mcd$Close))
carf$y = diff(log(carf$Close))
bay$y = diff(log(bay$Close))
cipla$y = diff(log(cipla$Close))
brk$y = diff(log(brk$Close))
jpm$y = diff(log(jpm$Close))
```

```{r, echo=FALSE, fig.width=10, fig.height=8, fig.align='center', out.width = '200%', out.height = '400%', dpi = 300}
# Now, let's visualize the closing prices of the chosen stocks along with their relative log-returns
plot(cbind(eni$Close,smh$Close,aapl$Close,micro$Close,mcd$Close,carf$Close,bay$Close,cipla$Close,brk$Close,jpm$Close), main = 'Stocks closing prices', xlab = 'Date', ylab = c('ENI','SMH','AAPL','MICRO','MCD','CARF','BAY','CIPLA','BRK','JPM'), cex.main = 2)
plot(cbind(eni$y,smh$y,aapl$y,micro$y,mcd$y,carf$y,bay$y,cipla$y,brk$y,jpm$y), main = 'Stocks relative log-returns', xlab = 'Date', ylab = c('ENI','SMH','AAPL','MICRO','MCD','CARF','BAY','CIPLA','BRK','JPM'), cex.main = 2)
```

```{r, echo=TRUE}
# PREPROCESSING PART 2
# Create a dataframe of the relative log-returns for each stock
# We've decided to drop the first value as the relative log-return can't be calculated for the first day
eni <- as.data.frame(eni$y[-1])
eni$dates <- rownames(eni)
smh <- as.data.frame(smh$y[-1])
smh$dates <- rownames(smh)
aapl <- as.data.frame(aapl$y[-1])
aapl$dates <- rownames(aapl)
micro <- as.data.frame(micro$y[-1])
micro$dates <- rownames(micro)
mcd <- as.data.frame(mcd$y[-1])
mcd$dates <- rownames(mcd)
carf <- as.data.frame(carf$y[-1])
carf$dates <- rownames(carf)
bay <- as.data.frame(bay$y[-1])
bay$dates <- rownames(bay)
cipla <- as.data.frame(cipla$y[-1])
cipla$dates <- rownames(cipla)
brk <- as.data.frame(brk$y[-1])
brk$dates <- rownames(brk)
jpm <- as.data.frame(jpm$y[-1])
jpm$dates <- rownames(jpm)

# Join the datasets together to get a data matrix of their relative log-returns
df <- plyr::join_all(list(eni,smh,aapl,micro,mcd,carf,bay,cipla,brk,jpm), by = 'dates', type = 'full')
df <- subset(df, select = -c(dates))
colnames(df) <- c('eni','smh','aapl','micro','mcd','carf','bay','cipla','brk','jpm')

# Print the correlation matrix between stocks
corr_mat <- cor(df)
# WIP - Prettify it
knitr::kable(corr_mat ,  caption = "Correlation Matrix")
```

WIP - SEARCH HOW TO GET NORMAL CI FOR PEARSON'S CORRELATION IN R, SELECT A THRESHOLD AND BUILD THE GRAPH
WIP - PLOT THE GRAPH WITH IGRAPHR AND COMMENT