---
title: "Homework 2 - Group 46"
output: html_document
date: "2023-12-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# A/B tests and Confidence Sequences

BRIEFLY: It's common practice in web-marketing to run large-scale randomized experiments called **A/B tests**. They consist in redirecting the traffic whether to one of two different versions of a website/video/media. Several data is registered, such as the *Click-Through Rate*, the *Time spent on-site*, etc.  
This permits to test if one of the two versions is more effective than the other.  
These experiments are **sequential**, meaning that the sample size is not fixed in advance: sampling is stopped in accordance with a pre-defined **stopping rule** (as soon as significant results are observed). So, using inferential methods that assume a fixed sample size is a **really bad idea** because it leads to p-hacking (you can find more informations inside the article in the repository).  
To be more specific, using a **Confidence Interval** to estimate a population parameter from a sequence of IID data (**observed sequentially**) may be restrictive: the sample size must be fixed beforehand and it is impossible to know *a priori* if the selected sample size will be large enough to detect some signal of interest.  
Instead, **Confidence Sequences** are more suitable: they can be updated whenever new data become available (sample size is chosen data-adaptively), without affecting the chosen coverage level. This makes possible to continuously monitor and adaptively stop (even at *data-dependent* times) the experiment. 
A **Confidence Sequence** is a sequence of confidence intervals such that each of them at time $t$ (so, after observing the $t^{th}$ unit) has got a coverage probability of $1-\alpha$.  
Given a sequence of IID observations from a 1-sub-Gaussian distribution (a distribution which tails decay at least as fast as those of a Gaussian distribution) the $1-\alpha$ confidence sequence for the population mean $\mu$ is:
\begin{align}
\frac{\sum_{i=1}^t X_i}{t} \pm 1.7 \cdot \sqrt{\frac{\log(\log(2t))+0.72\log(10.4/\alpha)}{t}}
\end{align}
Since this specific confidence sequence is *non-asymptotic*, it requires strong assumptions on the distribution of the data. So, the following *asymptotic* confidence sequence is introduced (suitable when only the asymptotic distribution of the sample mean is 1-Sub-Gaussian):
\begin{align}
\frac{\sum_{i=1}^t X_i}{t} \pm \hat{\sigma}_t \cdot 1.7 \cdot \sqrt{\frac{\log(\log(2t))+0.72\log(10.4/\alpha)}{t}}
\end{align}
Our job is to design a simulation to **compare the performance of the two introduced CSs with the classic CIs** (like Normal, Hoeffding, etc.). We must pick:

- one or two suitable populations to sample from
- the simulation size
- the target level $\alpha$
- the minimum and maximum sample sizes ($t_{min}$ and $t_{max}$)
- some metrics, like the *Cumulative Miscoverage Probability* and the *Running Interval Length*

Comments and visualizations are mandatory.  
Our second task consists of designing an experiment entitled **"Environmental Sound Levels and Mood"**, which investigates the relationship between the sound levels and the mood of a person over an extended period of time (so it's sequential). We must describe in details:

- how to organize the data collection
- what kind of tools/phone app/other stuff we would use
- what kind of intervention we may try to check the effect on the person
- the population parameter to monitor/estimate over time (and how CSs are employed in this)

## Exercise 1: Design a simulation to compare CSs and CIs

IDEAS AND METHODS ARE TAKEN FROM VARIOUS ARTICLES AND BLOGS, BUT MAINLY FROM [THIS](https://engineering.atspotify.com/2023/03/choosing-sequential-testing-framework-comparisons-and-discussions/) SPOTIFY SERIES OF ARTICLES ON SEQUENTIAL TESTING.  
YOU CAN FIND USEFUL CODE [HERE](https://github.com/MSchultzberg/SeqTestBlog/blob/main/20230321_sim_blogg.R)

The idea is to sample iteratively (until a number between t_min and t_max is reached) a random observation from two populations following a sub-gaussian distribution (maybe one gaussian and one ...). At each unit sampled the correspondent CSs and CIs are computed and informations about them are stored (like the bounds and the width). Also some others measures are calculated.  
This approach assumes that data comes in a continuous stream, instead of coming in batches. Plots will help to check how many times the value of the true parameter is included in the intervals.

```{r, echo=TRUE}
# FIRST CELL OF CODE
# Calculate the normal CI (which is asymptotic if the distribution is asymptotically normal, like the maximum likelihood estimator distribution)
ci_norm <- function(theta_ml, a, se){
  low <- theta_ml - qnorm(1-(a/2))*se
  up <- theta_ml + qnorm(1-(a/2))*se
  return(c(lower = low, upper = up))
}

# Calculate the Hoeffding CI
ci_hoef <- function(){
  low <- 
  up <- 
  return(c(lower = low, upper = up))
}

# Calculate Sub-Gaussian CS
cs_subnorm <- function(mu_est, a, csize){
  low <- mu_est - 1.7*sqrt((log(log(2*csize))+0.72*log(10.4/a))/csize)
  up <- mu_est + 1.7*sqrt((log(log(2*csize))+0.72*log(10.4/a))/csize)
  return(c(lower = low, upper = up))
}

# Calculate Asymptotic Sub-Gaussian CS (which is asymptotic if the distribution is asymptotically normal, like the maximum likelihood estimator distribution)
cs_asym_subnorm <- function(mu_est, var_est, a, csize){
  low <- mu_est - 1.7*var_est*sqrt((log(log(2*csize))+0.72*log(10.4/a))/csize)
  up <- mu_est + 1.7*var_est*sqrt((log(log(2*csize))+0.72*log(10.4/a))/csize)
  return(c(lower = low, upper = up))
}


# ---- SIMULATION ----
M <- 10000 # arbitrary
t_min <- 30 # arbitrary
t_max <- 500 # arbitrary
true_mu <- 2 # arbitrary
true_sd <- 1 # arbitrary
alpha <- 0.05 # arbitrary

# Initialize vectors to store sample and calculated CIs and CSs
samp <- c()
ci_norm_lows <- c()
ci_norm_ups <- c()
cs_lows <- c()
cs_ups <- c()
# Main loop
# REMEMBER TO ADD ONE MORE LOOP
set.seed(4545)
for (t in 1:t_max){
  samp <- c(samp, rnorm(1, mean = true_mu, sd = true_sd))
  if (t>=t_min){
    ci_norm_est <- ci_norm(mean(samp), alpha, sqrt(var(samp)/t))
    cs_est <- cs_subnorm(mean(samp), alpha, t)
    ci_norm_lows <- c(ci_norm_lows, ci_norm_est['lower'])
    ci_norm_ups <- c(ci_norm_ups, ci_norm_est['upper'])
    cs_lows <- c(cs_lows, cs_est['lower'])
    cs_ups <- c(cs_ups, cs_est['upper'])
  }
}
```

```{r, echo=FALSE}
# FIRST PLOT
plot(t_min:t_max, ci_norm_lows, type = 'l')
lines(t_min:t_max, ci_norm_ups, type = 'l')
lines(t_min:t_max, cs_lows, type = 'l', col = 'red')
lines(t_min:t_max, cs_ups, type = 'l', col = 'red')
```
