---
title: "Homework 2 - Group 46"
output: html_document
date: "2023-12-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# A/B tests

BRIEFLY: It's common practice in web-marketing to run large-scale randomized experiments called **A/B tests**. They consist in redirecting the traffic whether to one of two different versions of a website/video/media. Several data is registered, such as the *Click-Through Rate*, the *Time spent on-site*, etc.  
This permits to test if one of the two versions is more effective than the other.  
These experiments are **sequential**, meaning that the sample size is not fixed in advance: sampling is stopped in accordance with a pre-defined **stopping rule** (as soon as significant results are observed). So, using inferential methods that assume a fixed sample size is a **really bad idea** because it leads to p-hacking (you can find more informations inside the article in the repository).  
To be more specific, using a **Confidence Interval** to estimate a population parameter from a sequence of IID data (**observed sequentially**) may be restrictive: the sample size must be fixed beforehand and it is impossible to know *a priori* if the selected sample size will be large enough to detect some signal of interest.  
Instead, **Confidence Sequences** are more suitable: they can be updated whenever new data become available (sample size is chosen data-adaptively), without affecting the chosen coverage level. This makes possible to continuously monitor and adaptively stop (even at *data-dependent* times) the experiment. 
A **Confidence Sequence** is a sequence of confidence intervals such that each of them at time $t$ (so, after observing the $t^{th}$ unit) has got a coverage probability of $1-\alpha$.  
Given a sequence of IID observations from a 1-sub-Gaussian distribution (a distribution which tails decay at least as fast as those of a Gaussian distribution) the $1-\alpha$ confidence sequence for the population mean $\mu$ is:
\begin{align}
\frac{\sum_{i=1}^t X_i}{t} \pm 1.7 \cdot \sqrt{\frac{\log(\log(2t))+0.72\log(10.4/\alpha)}{t}}
\end{align}
Since this specific confidence sequence is *non-asymptotic*, it requires strong assumptions on the distribution of the data. So, the following *asymptotic* confidence sequence is introduced (suitable for data whose asymptotic distribution is 1-Sub-Gaussian):
\begin{align}
\frac{\sum_{i=1}^t X_i}{t} \pm \hat{\sigma}_t \cdot 1.7 \cdot \sqrt{\frac{\log(\log(2t))+0.72\log(10.4/\alpha)}{t}}
\end{align}
Our job is to design a simulation to **compare the performance of the two introduced CSs with the classic CIs** (like Normal, Hoeffding, etc.). We must pick:

- one or two suitable populations to sample from
- the simulation size
- the target level $\alpha$
- the minimum and maximum sample sizes ($t_{min}$ and $t_{max}$)
- some metrics, like the *Cumulative Miscoverage Probability* and the *Running Interval Length*

Comments and visualizations are mandatory.  
Our second task consists of designing an experiment entitled **"Environmental Sound Levels and Mood"**, which investigates the relationship between the sound levels and the mood of a person over an extended period of time (so it's sequential). We must describe in details:

- how to organize the data collection
- what kind of tools/phone app/other stuff we would use
- what kind of intervention we may try to check the effect on the person
- the population parameter to monitor/estimate over time (and how CSs are employed in this)


```{r, echo=TRUE}
# FIRST CELL OF CODE
```

```{r, echo=FALSE}
# FIRST PLOT
```
